---
title: "Lab 1"
author: "Veronica Weber"
date: "1/9/2020"
output: html_document
# Here - used to direct R to gather data sets from different data pathways
# Mutate - to change column asthetics while keeping the same column name (for instance changing the all caps text to lowercase)
# String (str_to_lower) - commands the text to change from upper to lower case
# ? (package) - use in the console - lets you find out what each package does, provides info
# Parse_Number - removes $ symbol from data cells
# String (str_remove) remove a string's pattern (keyword) from data cells
# String (str_detect) search for a pattern (keyword) and isolate data cells that match that pattern
# Separate - seperate data cells into different columns based on character/punctuation
# Summarize - creates a summary table and also takes the group of variables selected and reduces them down to a single value based on what I choosed (total, mean, etc.)
# GGsave - export a finished ggplot graph into specific folders 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)

```

Read in noaa_fisheries.csv data

```{r}
us_landings <- read_csv(here("data", "noaa_fisheries.csv"))
```

Make a tidy version of the data frame
```{r}
landings_tidy <- us_landings %>% 
  janitor::clean_names() %>% 
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_num = parse_number(dollars_usd))
```

Let's just get some information for salmon:

```{r}
salmon_landings <- landings_tidy %>% 
  mutate(afs_clean = str_remove(afs_name, pattern = "aggregate")) %>% 
  filter(str_detect(afs_clean, pattern = "salmon")) %>% 
  separate(afs_clean, into = c("group", "species"), sep = ",")
```

Find some grouped summary data:

Find annual total US landings and dollar value (summing across all states) for each TYPE of salmon using 'group_by' + 'summarize'

```{r}
salmon_summary <- salmon_landings %>% 
  group_by(year, species) %>% 
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_num)
  )
```


Create a graph for this info

```{r}
salmon_landings_graph <- ggplot(data = salmon_summary,
                                aes(x = year, y = tot_landings)) +
  geom_line(aes(color = species)) +
  theme_minimal()

salmon_landings_graph

```

Now we'll export our graph with ggsave

```{r}
ggsave(plot = salmon_landings_graph,
       here("figures", "us_salmon_vw.png"),
       height = 5,
       width = 8)
```

Let's make a nice table 
```{r}
salmon_first_5 <- salmon_summary %>% 
  head(5)
```

use kable to create a nice clean summary table
```{r}
kable(salmon_first_5) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
 ```

